---
title: "Tidy Tuesday 2019-51"
author: "James Wade"
slug: "tidy-tuesday-2019-51"
date: 2019-12-12
categories: ["R"]
tags: ["R Markdown", "plot", "data viz", "TidyTuesday"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here is my entry for the 51st [#TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-12-17) if 2019: Adoptable dogs.

## Import data

This week, I download the csv files provided in the GitHub. I'd like to focus more on the visualization than the cleaning. The data comes from [PetFinder](petfinder.org) via their [API](https://www.petfinder.com/developers/). Please see the post from [The Pudding](https://github.com/the-pudding/data/blob/master/dog-shelters/README.md) for additional details. 

```{r import_data, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
library(ggtext)
library(geofacet)
library(sf)
library(rnaturalearth)
library(viridis)

moves <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-12-17/dog_moves.csv')
descriptions <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-12-17/dog_descriptions.csv')
travel <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-12-17/dog_travel.csv')
```

Since the cleaning has already been done for me, I'll skip to visulizaitons

## Moves

I live in Michigan, and we hear a lot about how often dogs are transported here for adoption. I'd like to create a visual of flow of dogs between states. Starting with a simple bar chart, we can see net movement of dogs. Note: it looks like Texas has a reporting issue, so I'll remove it.

```{r move, message=FALSE, warning=FALSE}
states <- tibble(states_abb = state.abb,
                 states_name = state.name)

m <- moves %>% 
  filter(inUS == TRUE) %>% 
  mutate(imported = case_when(is.na(imported) ~ 0,
                              TRUE ~ imported),
         exported = case_when(is.na(exported) ~ 0,
                              TRUE ~ exported),
         net_flow = (imported - exported) / total * 100) %>% 
  left_join(states, by = c("location" = "states_name")) %>%
  filter(location != "Texas" & !is.na(states_abb))

p_net <- ggplot(m, aes(x = reorder(location, net_flow), y = net_flow, fill = net_flow > 0)) +
  labs(x = "State", y = "Percent of Adoptable Dogs") +
  scale_fill_manual(values = c("#ef8a62", "#67a9cf")) +
  geom_bar(stat = "identity") +
  annotate("text", label = "Dog Importer", x = 5, y = 25, color = "#67a9cf", size = 5) +
  annotate("text", label = "Dog Exporter", x = 46, y = -25, color = "#ef8a62", size = 5) +
  annotate("rect", xmin=0, xmax=51, ymin=0, ymax=30, alpha = 0.3, fill = "#67a9cf") +
  annotate("rect", xmin=0, xmax=51, ymax=0, ymin=-30, alpha = 0.3, fill = "#ef8a62") +
  theme_minimal() + 
  labs(title = "**<span style='color:#67a9cf'>Imports</span> and <span style='color:#ef8a62'>Exports</span> of Dogs by State**",
       subtitle = "Source: PetFinder | Texas Excluded (Suspected Outlier)",
       caption = "#TidyTuesday | @JamesHWade") +
  theme(plot.title = element_markdown(hjust = 0.5, size = 20),
        plot.subtitle = element_markdown(hjust = 0.5, size = 10),
        legend.position = "none",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1))
p_net

ggsave(filename = "51_geombar.png", plot = p_net,
       dpi = "retina", width = 8, height = 5)

m2 <- moves %>% 
  filter(inUS == TRUE) %>% 
  mutate(imported = case_when(is.na(imported) ~ 0,
                              TRUE ~ imported),
         exported = case_when(is.na(exported) ~ 0,
                              TRUE ~ exported),
         import_percent = imported / total * 100,
         export_percent = -exported / total * 100,
         net_flow = (imported - exported) / total * 100) %>% 
  pivot_longer(c(import_percent, export_percent, net_flow)) %>% 
  filter(location != "Texas") # removing presumed outlier

ggplot(m2, aes(x = reorder(location, value), y = value, fill = name)) + geom_col(position = "dodge") +
  theme_minimal() + 
  theme(legend.position = "none") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```

Just because I can and how easy it is, I'll make a plotly graph from that same plot.

```{r move_plotly, message=FALSE, warning=FALSE}
plotly::ggplotly(p_net)
```

### Making Maps with `geom_sf`

To be honest, this took me way too long to get this plot to work, but learning is the whole point of #TidyTuesday, right? Below is the same data as above but shown on a map.

```{r move_map, message=FALSE, warning=FALSE}
usa <- ne_states(country = "United States of America", returnclass = 'sf')

pet_data <- usa %>% left_join(m, by = c('name' = 'location'))

mainland <- ggplot(data = pet_data) +
     geom_sf(aes(fill = net_flow), size = 0.2) +
     coord_sf(crs = st_crs(2163), xlim = c(-2500000, 2500000), ylim = c(-2300000, 
         730000)) + theme_map() +
  scale_fill_gradient2(low = "#ef8a62", high = "#67a9cf") +
  labs(title = "**<span style='color:#67a9cf'>Imports</span> and <span style='color:#ef8a62'>Exports</span> of Dogs by State**",
       subtitle = "Source: PetFinder | Texas Excluded (Suspected Outlier)",
       caption = "#TidyTuesday | @JamesHWade") +
  theme(plot.title = element_markdown(hjust = 0.5, size = 20),
        plot.subtitle = element_markdown(hjust = 0.5, size = 10)) +
  labs(fill = "Net Flow (%)", color = "Net Flow (%)", parse = TRUE)

mainland

ggsave(filename = "51_geomsf.png", plot = mainland,
       dpi = "retina")
```

And, surprise! We can make a plotly graph again.

```{r map_plotly, message=FALSE, warning=FALSE}
plotly::ggplotly(mainland)
```


### `facet_geo` - My new Favorite Plot

I've found a few plots on twitter using `facet_geo`, and I just had to give it a try. I took a lot of the code from [Thomas Mock](https://twitter.com/thomas_mock/status/1206641246351048705), but others put out similar looking plots.

```{r geofacet, message=FALSE, warning=FALSE}
p_geofacet <- ggplot(m, aes(x = "Dogs", y = net_flow, fill = net_flow > 0)) +
  scale_fill_manual(values = c("#ef8a62", "#67a9cf")) +
  geom_bar(stat = "identity") +
  theme_minimal() + 
  theme(legend.position = "none",
        axis.text = element_blank(),
        axis.title = element_blank(),
        plot.title = element_markdown(hjust = 0.5, size = 20),
        plot.subtitle = element_markdown(hjust = 0.5, size = 10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  labs(title = "**<span style='color:#67a9cf'>Imports</span> and <span style='color:#ef8a62'>Exports</span> of Dogs by State**",
       subtitle = "Source: PetFinder | Texas Excluded (Suspected Outlier)",
       caption = "#TidyTuesday | @JamesHWade") +
  facet_geo(~states_abb, grid = "us_state_grid2")

p_geofacet

ggsave(filename = "51_geofacet.png", plot = p_geofacet,
       dpi = "retina")
```