---
title: Tidy Tuesday Extra - Bring Your Own Data
author: James Wade
date: '2019-12-31'
slug: tidy-tuesday-extra-bring-your-own-data
categories:
  - R
  - government
  - spending
tags:
  - data viz
  - plot
  - R Markdown
  - TidyTuesday
---

For this 53rd Tuesday in 2019, [#TidyTuesday](https://github.com/rfordatascience/tidytuesday) is a byod (bring-your-own-data) special. For this week, I'll take a look at one of the weeks I missed from earlier in 2019. I'm only a few years removed from my grad school days, so government spending on R&D is an interest of mine. Week 7 it is: Federal R&D Spending. This is a very short post, but I have family in town, so I don't wan tot go too far down a rabbit hole.

## Data Import

For import, I'll use the `{tidytuesdayR}` package - quick and easy.

```{r}
# import data using the {tidytuesdayR} package
dat <- tidytuesdayR::tt_load(2019, week = 7)

# assign each element in list to variable
fed_tot <- dat$fed_r_d_spending
climate <- dat$climate_spending
energy <- dat$energy_spending
```

To get a sense of relavtive changes by year, I'll add a few columns with `mutate`. These basic tranformations enable comparisons to total and discresionary spending.

```{r fed_spending, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)

# summarise spending by department as percentage of GDP, discretionary, & total spending
fed_tot <- fed_tot %>%
  group_by(department) %>% 
  mutate(rd_percent_tot = rd_budget / total_outlays * 100,
         rd_percent_disc = rd_budget / discretionary_outlays * 100,
         rd_vs_gdp = rd_budget / gdp * 100) %>% 
  arrange(rd_percent_tot) %>% 
  ungroup()

# summarise by defense and non
```


```{r fed_spending_plots, fig.height=6, fig.width=6, fig.align="center", message=FALSE, warning=FALSE}
p1_dollars <- ggplot(fed_tot) + 
  geom_area(aes(x = year, y = rd_budget / 1e9, fill = reorder(department, -rd_percent_tot))) + 
  theme_cowplot() +
  theme(legend.position = "bottom") +
  labs(x = "", y = "R&D Spending \n(Billions of Dollars)", fill = "Agency",
       title = "US Federal R&D Spending over Time",
       subtitle = "Source: AAAS | 1976-2017",
       caption = "#TidyTuesday | @JamesHWade") +
  scale_fill_viridis_d()

p1_dollars

p1_percent_tot <- ggplot(fed_tot) + 
  geom_area(aes(x = year, y = rd_percent_tot, fill = reorder(department, -rd_percent_tot))) + 
  theme_cowplot() +
  theme(legend.position = "bottom") +
  labs(x = "", y = "R&D Spending \n(Percent of Total Spending)", fill = "Agency",
       title = "US Federal R&D Spending as Percentage of Total",
       subtitle = "Source: AAAS | 1976-2017",
       caption = "#TidyTuesday | @JamesHWade") +
  scale_fill_viridis_d()

p1_percent_tot

p1_percent_gdp <- ggplot(fed_tot) + 
  geom_area(aes(x = year, y = rd_vs_gdp, fill = reorder(department, -rd_percent_tot))) + 
  theme_cowplot() +
  theme(legend.position = "bottom") +
  labs(x = "", y = "R&D Spending \n(Percentage of GDP)", fill = "Agency",
       title = "US Federal R&D Spending as Percentage of GDP",
       subtitle = "Source: AAAS | 1976-2017",
       caption = "#TidyTuesday | @JamesHWade") +
  scale_fill_viridis_d()

p1_percent_gdp

# uncomment ot save
ggsave(p1_dollars, filename = "total_dollars.png", width = 6, height = 6, dpi = "retina")
ggsave(p1_percent_tot, filename = "total_percent.png", width = 6, height = 6, dpi = "retina")
ggsave(p1_percent_gdp, filename = "total_gdp.png", width = 6, height = 6, dpi = "retina")
```

## Defense vs Non-Defense

To get a sense of defense versus non-defense spenidng, I've re-color the plots as either defense or non-defense.

```{r def_vs_non_line, fig.height=6, fig.width=6, fig.align="center"}
p2_dollars <- ggplot(fed_tot) + 
  geom_area(aes(x = year, 
                y = rd_budget / 1e9, 
                fill = department == "DOD", 
                group = reorder(department, -rd_percent_tot))) + 
  theme_cowplot() +
  scale_fill_manual(values = c("#91bfdb", "#fc8d59"), 
                      labels = c("Non-Defense", "Defense")) +
  theme(legend.position = "bottom") +
  labs(x = "", y = "R&D Spending \n(Billions of Dollars)", fill = "Agency",
       title = "US Federal R&D Spending over Time",
       subtitle = "Source: AAAS | 1976-2017",
       caption = "#TidyTuesday | @JamesHWade")

p2_dollars

p2_percent_tot <- ggplot(fed_tot) + 
  geom_area(aes(x = year,
                y = rd_percent_tot, 
                fill = department == "DOD", 
                group = reorder(department, -rd_percent_tot))) + 
  theme_cowplot() +
  scale_fill_manual(values = c("#91bfdb", "#fc8d59"), 
                      labels = c("Non-Defense", "Defense")) +
  theme(legend.position = "bottom") +
  labs(x = "", y = "R&D Spending \n(Percentage of Total Spending)", fill = "Agency",
       title = "US Federal R&D Spending as Percentage of Total",
       subtitle = "Source: AAAS | 1976-2017",
       caption = "#TidyTuesday | @JamesHWade")

p2_percent_tot

p2_percent_gdp <- ggplot(fed_tot) + 
  geom_area(aes(x = year, 
                y = rd_vs_gdp, 
                fill = department == "DOD", 
                group = reorder(department, -rd_percent_tot))) + 
  theme_cowplot() +
  scale_fill_manual(values = c("#91bfdb", "#fc8d59"), 
                      labels = c("Non-Defense", "Defense")) +
  theme(legend.position = "bottom") +
  labs(x = "", y = "R&D Spending \n(Percentage of GDP)", fill = "Agency",
       title = "US Federal R&D Spending as Percentage of GDP",
       subtitle = "Source: AAAS | 1976-2017",
       caption = "#TidyTuesday | @JamesHWade")

p2_percent_gdp

# uncomment to save
ggsave(p2_dollars, filename = "total_dollars_2.png", width = 6, height = 6, dpi = "retina")
ggsave(p2_percent_tot, filename = "total_percent_2.png", width = 6, height = 6, dpi = "retina")
ggsave(p2_percent_gdp, filename = "total_gdp_2.png", width = 6, height = 6, dpi = "retina")
```