---
title: "Tidy Tuesday 2019-50"
author: "James Wade"
date: 2019-12-12
categories: ["R"]
tags: ["R Markdown", "plot", "data viz", "TidyTuesday"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

Here is my entry for [#TidyTuesday](https://github.com/rfordatascience/tidytuesday) from 2019-50: 'You can make it in R.'

## Import the Data

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(gganimate)
library(waffle)

d <- tidytuesdayR::tt_load(2019, week = 50)

diseases <- d$diseases
murders <- d$international_murders
guns <- d$gun_murders
regents <- d$nyc_regents
```

# Part 1: Diseases

The first plot was inspired by an [interactive graphic](http://graphics.wsj.com/infectious-diseases-and-vaccines/) from WSJ. The data for these visuals comes from the [Tycho Project](https://www.tycho.pitt.edu).

To start, let's add a column that provides a *rate* of disease incidence. The metrics will be in using of cases per 100,000 people per year.

```{r}
disease_2 <- diseases %>% 
  group_by(year, state, disease) %>% 
  mutate(rate = count / population * 10000 * 52 / weeks_reporting) %>% 
  ungroup()
```

## Measles Plot with Vaccine

The graphic from the blog post uses Measles. This is a particularly impactful graphic emphasizing the impact of the measles vaccine on the rate of measles occurence across US states. Rather than improve upon this already impressive graph, I chose to animate it instead. This uses the gganimate package. This animation only require an additional 3 lines of code. So fancy for such little effort!

```{r}
jet.colors <- colorRampPalette(c("#F0FFFF", "cyan", "#007FFF", "yellow", 
                                 "#FFBF00", "orange", "red", "#7F0000"), bias = 2.25)


p_anim <- ggplot(disease_2 %>% filter(disease == "Measles"), aes(x = year, y = state, fill = rate)) + 
  geom_tile(color = "white", size = 0.35) +
  scale_x_continuous(expand = c(0,0)) +
  scale_fill_gradientn(colors = jet.colors(16), na.value = 'white') +
  theme_minimal() +
  transition_time(year) + 
  shadow_mark() +
  annotate(geom = "text", x = 1963.5, y = 50.5, label = "Vaccine introduced", size = 5, hjust = 0) +
  ggtitle("Incidence of Meases (USA) \n1928-2011") +
  geom_vline(xintercept = 1963, col = "black") +
  labs(x = "Year", y = "State", fill = "Rate\n(Cases per 100,000)")

# uncomment to save animation
# anim_save(p_anim, width = 1000, height = 600, filename = "measles.gif")

animate(p_anim, width = 1000, height = 600)
```

During the [2019-49 #TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-12-03), I saw some [waffle plots on Twitter](https://twitter.com/privlko/status/1201918825941360640). These had me intrigued, so I decided to play around with `geom_waffle` from the [waffle](https://cinc.rud.is/web/packages/waffle/index.html) library.

For this visual, I filtered for a single state (Georgia) and limited the time frame from 1976 to 1995. With the waffle plot, you can see that there were upticks in measles cases in 1977, 1979, 1980, and 1990. From some quick web searching, there does appear to have been an outbreak of measles in GA in 1990.

```{r}
disease_3 <- disease_2 %>% 
  filter(state == "Georgia" & year %in% c(1976:1995))

p_waffle <- ggplot(disease_3) +
  geom_waffle(aes(fill = disease, values = count), 
              color = "white", n_rows = 50, flip = TRUE) +
  facet_wrap(~year) +
  coord_equal() +
  ggtitle("Measles Outbreak in Geogia \n(1976-1995)") +
  theme_classic()

p_waffle
```

As a comparison, I also made a line graph with the same data.

```{r}
p_line <- ggplot(disease_3) +
  geom_line(aes(x = year,color = disease, y = count)) +
  theme_classic()

p_line
```